### 2021-12-14 thesis/tex/conclusion.tex
update

\chapter{Conclusion}\label{ch:conc}
In this thesis, we have explored the potential of explaining new physics using detectors handled by the IceCube collaboration.
We started off by introducing the Standard Model in Sec.~\ref{ch:SM}, and noted that it has several shortcomings. 
We mentioned the Super-Kamiokande experiment, and that their 1998 result found no other explanation than $\ne \to \nm$ mixing~\cite{sk1998}, in direct violation of the Standard Model. 
We saw that, by adding a right-handed neutrino field, neutrino masses are generated by the Higgs mechanism, thus enabling neutrino oscillations to occur. 
In Sec.~\ref{ch:oscillation}, we saw how the neutrino mixing matrix could be constructed, and derived formulas describing neutrino oscillations in the two- and three-neutrino pictures. Adding a term for matter potential,
we then arrived at a Hamiltonian that we used in the time-dependent Schrödinger equation to solve for the oscillation probabilities in matter.

@@

We then put the sterile neutrino aside, and introduced Non-Standard Interactions (NSI) in Sec.\ref{sec:nsiTheory} as an possible example of new physics beyond the Standard Model.
We then studied the effects of NSI on the neutrino oscillation probabilities and saw that NSI effects other than those stemming from $\emt$ are not apparent in the \si{\TeV} region.
Moreover, we saw in $\Pme$ that we could expect a correlation between $\eem$ and $\eet$ in the single-digit \si{\GeV} range.
We also saw that $\emt$ effects were visible on probability level in both the \si{\GeV} and \si{\TeV} regions.

First, we described our method in testing the sterile neutrino hypothesis (3+1). We simulated the 86 string IceCube detector with track events, explained in Sec.\ref{ch:ICmethod}.
After simulating event counts in Sec.~\ref{sec:sterileResults} for the $3+1$ hypothesis, we were able to find best-fit parameters of 
$\dm[41] = \SI{0.01}{\eV^2}$ and $\theta_{24} = 0.67$ ($\sin^22\theta_{24} = 0.95$) at 
a p-value of $20\%$. Even though we found a best-fit for two parameters of the sterile neutrino, they are not statistically significant. Hence, we found no evidence of a sterile neutrino in IceCube data.
We also investigated the case when $\theta_{24} = \theta_{34}$, which produced an identical contour as in the case with $\theta_{34}=0$. From this observation, 
we concluded that IceCube is currently insensitive to any impact of $\theta_{34}$, with track events alone. However,
when opening up the $\tau$ channel by allowing $\theta_{34} \neq 0$, it is possible that with an inclusion of cascade events from $\nt$, IceCube could be able to better observe 
an impact of $\theta_{34}$. With these results, we then completely left the sterile neutrino behind.

@@

In Sec.~\ref{ch:DCmethod}, we moved on and explained our method of simulating a second detector, DeepCore. Armed with simulations possible for this detector, we then continued to explore the possibility of NSI, this time including both track and cascade events and a lower energy region.
Using the $\chi^2$ test defined in Eq.~\ref{eq:chisq_NSI} and after marginalizing $\dm[31]$ and $\theta_{23}$ out, we found the following constraints on the NSI parameters at 90\% confidence level:
\begin{align}
    -0.054 <&\, \ett < 0.067 \nonumber \\
    -0.029 <&\, \emt < 0.0070 \nonumber \\
    -0.12 <&\, \eem < 0.15 \nonumber \\
    -0.084 <&\, \eet < 0.15\ \quad (\text{all at }90\% \text{ CL})\,.
 \end{align}
We compared these constraints to literature, and found that the bounds for $\eem$ and $\eet$ were more stringent than those obtained in~\cite{demidov}.
Then, we did a joint $\chi^2$ test using simulated events from both IceCube and DeepCore. At 90\% confidence level, the bound of $\emt$ improved to
\begin{align}
    -0.016 <&\, \emt < 0.0070\ \quad (90\% \text{ CL})\,.
 \end{align}

@@

We were then able to draw the conclusion that we we expect PINGU to successfully be able to observe the impact of NSI, and a joint analysis between PINGU and at least one of the other detectors under the IceCube collaboration
will be able to even further constrain the parameters. We can expect PINGU to be more sensitive to NSI effects than DeepCore. Moreover, we saw that when constraining $\eem$ and $\eet$ from the negative side, 
we can expect the result to be highly dependent on the systematic uncertainty of PINGU.

Finally, we simulated the event count in PINGU allowing $\emt$, $\eem$, and $\eet$ to vary, but setting $\ett=0$.
After and marginalizing out $\dm[31]$, $\theta_{23}$, and $\emt$, we saw that the effect of $\eem$ and $\eet$ on $\Pme$ had indeed propagated through to the event count,
manifesting as an anti-correlation between the $\eem$ and $\eet$ at both 90\% and $3\sigma$ confidence levels. No other pair of NSI parameters were observed to have 
a significant correlation.

>>>>

\chapter{Conclusion}\label{ch:conc}
In this thesis, we have explored the potential of explaining new physics using detectors handled by the IceCube collaboration.
We started off by introducing the Standard Model in Sec.~\ref{ch:SM}, and noted that it has several shortcomings. 
We mentioned the Super-Kamiokande experiment and that their 1998 result found no other explanation than $\ne \to \nm$ mixing~\cite{sk1998}, in direct violation of the Standard Model. 
We saw that, by adding a right-handed neutrino field, neutrino masses are generated by the Higgs mechanism, thus enabling neutrino oscillations to occur. 
In Sec.~\ref{ch:oscillation}, we saw how the neutrino mixing matrix could be constructed and derived formulas describing neutrino oscillations in the two- and three-neutrino pictures. Adding a term for matter potential,
we then arrived at a Hamiltonian that we used in the time-dependent Schrödinger equation to solve for the oscillation probabilities in matter.

@@

We then put the sterile neutrino aside and introduced Non-Standard Interactions (NSI) in Sec.\ref{sec:nsiTheory} as a possible example of new physics beyond the Standard Model.
We then studied the effects of NSI on the neutrino oscillation probabilities and saw that NSI effects other than those stemming from $\emt$ are not apparent in the \si{\TeV} region.
Moreover, we saw in $\Pme$ that we could expect a correlation between $\eem$ and $\eet$ in the single-digit \si{\GeV} range.
We also saw that $\emt$ effects were visible on probability level in both the \si{\GeV} and \si{\TeV} regions.

First, we described our method in testing the sterile neutrino hypothesis (3+1). We simulated the 86 string IceCube detector with track events, explained in Sec.\ref{ch:ICmethod}.
After simulating event counts in Sec.~\ref{sec:sterileResults} for the $3+1$ hypothesis, we were able to find best-fit parameters of 
$\dm[41] = \SI{0.01}{\eV^2}$ and $\theta_{24} = 0.67$ ($\sin^22\theta_{24} = 0.95$) at 
a p-value of $20\%$. Even though we found a best-fit for two parameters of the sterile neutrino, they are not statistically significant. Hence, we found no evidence of a sterile neutrino in IceCube data.
We also investigated the case when $\theta_{24} = \theta_{34}$, which produced an identical contour as in the case with $\theta_{34}=0$. From this observation, 
we concluded that IceCube is currently insensitive to any impact of $\theta_{34}$, with track events alone. However,
when opening up the $\tau$ channel by allowing $\theta_{34} \neq 0$, it is possible that with the inclusion of cascade events from $\nt$, IceCube could be able to better observe 
an impact of $\theta_{34}$. With these results, we then completely left the sterile neutrino behind.

@@

In Sec.~\ref{ch:DCmethod}, we moved on and explained our method of simulating a second detector, DeepCore. Armed with simulations possible for this detector, we then continued to explore the possibility of NSI, this time including both track and cascade events and a lower energy region.
Using the $\chi^2$ test defined in Eq.~\ref{eq:chisq_NSI} and after marginalizing $\dm[31]$ and $\theta_{23}$ out, we found the following constraints on the NSI parameters at 90\% confidence level:
\begin{align}
    -0.054 <&\, \ett < 0.067 \nonumber \\
    -0.029 <&\, \emt < 0.0070 \nonumber \\
    -0.12 <&\, \eem < 0.15 \nonumber \\
    -0.084 <&\, \eet < 0.15\ \quad (\text{all at }90\% \text{ CL})\,.
 \end{align}
We compared these constraints to literature and found that the bounds for $\eem$ and $\eet$ were more stringent than those obtained in~\cite{demidov}.
Then, we did a joint $\chi^2$ test using simulated events from both IceCube and DeepCore. At 90\% confidence level, the bound of $\emt$ improved to
\begin{align}
    -0.016 <&\, \emt < 0.0070\ \quad (90\% \text{ CL})\,.
 \end{align}

@@

We were then able to draw the conclusion that we expect PINGU to successfully be able to observe the impact of NSI. Moreover, a joint analysis between PINGU and at least one of the other detectors under the IceCube collaboration
will be able to constrain the parameters even further. We saw that can expect PINGU to be more sensitive to NSI effects than DeepCore. We were also able to see that when constraining $\eem$ and $\eet$ from the negative side, 
we can expect the result to be highly dependent on the systematic uncertainty of PINGU.

Finally, we simulated the event count in PINGU, allowing $\emt$, $\eem$, and $\eet$ to vary, but setting $\ett=0$.
After and marginalizing out $\dm[31]$, $\theta_{23}$, and $\emt$, we saw that the effect of $\eem$ and $\eet$ on $\Pme$ had indeed propagated through to the event count,
manifesting as an anti-correlation between the $\eem$ and $\eet$ at both 90\% and $3\sigma$ confidence levels. No other pair of NSI parameters were observed to have 
a significant correlation.

### 2021-12-14 thesis/tex/introduction.tex
update

\chapter{Introduction}
There is no reason for us to believe that our current knowledge in particle physics is complete.
The most successful framework -- the Standard Model -- has notable shortcomings. Neither does it
explain why there is something rather than nothing\footnote{In the Standard Model,
matter and antimatter come in pairs of one particle and one antiparticle. When interacting,
antimatter \emph{annihilates} the matter, resulting in pure light. Since the universe today contains matter,
there must have been more matter than antimatter at the Big Bang. This baryon asymmetry is unexplained in the Standard 
Model.}, nor does it give a candidate for dark matter\footnote{Dark matter is heavy matter which does not interact with electromagnetic radiation.
No such particle exists in the Standard Model.}.
To make matters even worse, three of the particles which are massless according to the Standard Model have been observed to have mass.
In this thesis, we focus on the said particles -- the neutrinos -- and propose extensions to the Standard Model related to them. 
We then compare the effects of these extensions with collected data to see if our new theory more closely describes Nature or not.

@@

Chapter~\ref{ch:theory} consists of two parts. In Section~\ref{sec:anomalies}, we introduce a new particle: the sterile neutrino. 
We present how neutrino oscillations are modified by this hypothesized fourth neutrino and how these modifications would appear in IceCube if the particle is present in Nature. 
In Section~\ref{sec:nsiTheory}, we completely set the sterile neutrino aside and instead turn to the interactions between neutrinos and the Earth. 
We again amend the Standard Model by considering exotic interactions from a higher energy theory, manifesting themselves as sub-leading modifications to the matter potential. 
Since these new interactions are not present in the Standard Model, we call them non-standard interactions (NSI).

>>>>

\chapter{Introduction}
There is no reason for us to believe that our current knowledge in particle physics is complete.
The most successful framework -- the Standard Model -- has notable shortcomings. Neither does it
explain why there is something rather than nothing\footnote{In the Standard Model,
matter and antimatter come in pairs of one particle and one antiparticle. When interacting,
antimatter \emph{annihilates} the matter, resulting in pure light. Since the universe today contains matter,
there must have been more matter than antimatter at the Big Bang. This baryon asymmetry is unexplained in the Standard 
Model.}, nor does it give a candidate for dark matter\footnote{Dark matter is heavy matter which does not interact with electromagnetic radiation.
No such particle exists in the Standard Model.}.
To make matters even worse, three of the particles which are massless according to the Standard Model have been observed to actually have mass.
In this thesis, we focus on the said particles -- the neutrinos -- and propose extensions to the Standard Model related to them. 
We then compare the effects of these extensions with collected data to see if our new theory more closely describes Nature or not.

@@

Chapter~\ref{ch:theory} consists of two parts. In Section~\ref{sec:anomalies}, we introduce a new particle: the sterile neutrino. 
We present how neutrino oscillations are modified by this hypothesized fourth neutrino and how these modifications would appear in IceCube if the particle is present in Nature. 
In Section~\ref{sec:nsiTheory}, we set the sterile neutrino aside altogether and instead turn to the interactions between neutrinos and the Earth. 
We again amend the Standard Model by considering exotic interactions from a higher energy theory, manifesting themselves as sub-leading modifications to the matter potential. 
Since these new interactions are not present in the Standard Model, we call them non-standard interactions (NSI).

### 2021-12-14 thesis/tex/old_abstract.tex
update

The failure of the Standard Model of particle physics to predict neutrino masses invites us to amend it. We have no reason to believe that 
the interactions currently described by the gauge group are the whole picture, nor should we expect that the number of fermions hitherto observed is correct.
In this thesis, we explore two amendments to the Standard Model, and examine whether either of these are consistent with measured data.

Firstly, we consider the sterile neutrino, which have no interactions described by the Standard Model. We examine the effect of this 
new fermion on the neutrino oscillation probabilities, and present what could be a detectable signal in the \si{\TeV} energy range.
Secondly, we consider interactions beyond the Standard Model, possibly stemming from a higher-order theory.
We show how the parameters of these Non-Standard Interactions (NSI) can modify the oscillation probabilities, and within which energy range we expect to be able to discern this signal.
We use data from and simulate event counts in two Cherenkov detectors: IceCube and DeepCore. Moreover, we generate data and simulate a proposed upgrade of the DeepCore detector: PINGU.

Using IceCube track events, we obtain best-fit values $\dm[41] = \SI{0.01}{\eV^2}$ and $\theta_{24} = 0.67$ for our sterile neutrino hypothesis at 
a p-value of $20\%$, which is not statistically significant. Hence, we found no evidence of a sterile neutrino in IceCube data. 
Moreover, were unable to distinguish a signal from $\theta_{34}$ in our IceCube simulation.
We obtain stringent bounds on the NSI parameters, and compare those to previous results in literature. We show that PINGU is expected to further narrow the bound 
on $\emt$, especially the considering a joint analysis together with IceCube and DeepCore. Finally, we see that an anti-correlation between 
$\eem$ and $\eet$ at probability level was propagated down to event level, which we expect to be observable by PINGU.

>>>>

The failure of the Standard Model of particle physics to predict neutrino masses forces us to amend it. We have no reason to believe that 
the interactions currently described by the gauge group are the whole picture, nor should we expect that the number of fermions hitherto observed is correct.
In this thesis, we explore two modifications to the Standard Model, and examine whether these amendments are consistent with measured data.
Firstly, we introduce the sterile neutrino, which have no interactions described by the Standard Model. We examine the effect of this 
new fermion on the neutrino oscillation probabilities, and present what would be a detectable signal in the \si{\TeV} range.
Secondly, we are introducing a new interaction which we call Non-Standard Interactions (NSI), possibly stemming from a higher-order theory. 
We show how the NSI parameters can modify the oscillation probabilities, and within which energy range we expect to be able to discern this signal.

We are collecting data from and simulating two Cherenkov detectors: IceCube and DeepCore. Moreover, we are presenting results which we 
can expect to see in a coming upgrade of the DeepCore detector: PINGU.

We obtain results indicating that the current 3+0 hypothesis is consistent with IceCube data, with best-fit values $\dm[41] = \SI{0.01}{\eV^2}$ and $\theta_{24} = 0.67$ at 
a p-value of $33\%$. We were unable to distinguish a signal from $\theta_{34}$.
We obtain stringent bounds on the NSI parameters, and compare those to previous results. We show that PINGU is expected to further narrow the bound 
on $\emt$, especially the considering a joint analysis together with IceCube and DeepCore. Finally, we see that an anti-correlation between 
$\eem$ and $\eet$ at probability level was propagated down to event level, which will be observable by PINGU.
